### 2022-10-12 Quarto_Simulation/index.qmd
Save simulation code.

---
title: "Simulation"
author: "Jeremy Selva"
format: html
---

# R Libaries used
```{r R packages}
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(purrr)
library(ggplot2)
library(epifitter)
library(DCVtestkit)

set.seed(123)
```

```{r Saturation Many Data}

simulate_saturation <- function(id) {

  mono_model <- sim_monomolecular(
  N = 200,     # total time units 
  y0 = 0.01,   # initial inoculum
  dt = 30,     #  interval between assessments in time units
  r = 0.0175,  #  apparent infection rate
  alpha = 0.15,# level of noise
  n = 1        # number of replicates
  )

  saturation_simulation <- tibble::tibble(
    x = mono_model$time,
    y = mono_model$random_y * sample(1:1000000, 1),
    `Sample Name` = as.character(mono_model$time))

  return(saturation_simulation)

}

saturation_simulated_database <-
  tibble::tibble(
    ID = paste("Saturation", 
               stringr::str_pad(1:200, width = 3, pad = "0")),
    `Dilution Batch Name` = "Saturation"

  ) |> 
  dplyr::mutate(
    dilution_data = purrr::map(
      .x = .data$ID,
      .f = simulate_saturation
      )
    ) |> 
  tidyr::unnest("dilution_data")

```

```{r}
# exp_model <- sim_exponential(
#     N = 200,    # total time units 
#     y0 = 1,  # initial inoculum
#     dt = 30,    #  interval between assessments in time units
#     r = 0.02,  #  apparent infection rate
#     #alpha = 0.01,# level of noise
#     alpha = 0.2, # level of noise
#     n = 1       # number of replicates
#     )
# 
# exp_plot = exp_model %>%
#   ggplot(aes(time, y)) +
#   geom_jitter(aes(time, random_y), size = 3, width = .1) +
#   geom_line(size = 1) +
#   labs(
#     title = "Exponential",
#     y = "Disease intensity",
#     x = "Time"
#   )
# exp_plot
```

```{r LOD Many Data}

simulate_lod <- function(id) {

  exp_model <- sim_exponential(
    N = 200,    # total time units 
    y0 = 1,  # initial inoculum
    dt = 30,    #  interval between assessments in time units
    r = 0.02,  #  apparent infection rate
    #alpha = 0.1, # level of noise
    alpha = 0.15, # level of noise
    n = 1       # number of replicates
    )

  lod_simulation <- tibble::tibble(
    x = exp_model$time,
    y = exp_model$random_y * sample(1:1000000, 1),
    `Sample Name` = as.character(exp_model$time))

  return(lod_simulation)

}

lod_simulated_database <-
  tibble::tibble(
    ID = paste("Limit Of Detection", 
               stringr::str_pad(1:200, width = 3, pad = "0")),
    `Dilution Batch Name` = "Limit Of Detection"

  ) |> 
  dplyr::mutate(
    dilution_data = purrr::map(
      .x = .data$ID,
      .f = simulate_lod
      )
    ) |> 
  tidyr::unnest("dilution_data")

```

```{r Linear Many Data}

simulate_lod <- function(id) {

  exp_model <- sim_exponential(
    N = 200,    # total time units 
    y0 = 1,  # initial inoculum
    dt = 30,    #  interval between assessments in time units
    r = 0.02,  #  apparent infection rate
    #alpha = 0.1, # level of noise
    alpha = 0.15, # level of noise
    n = 1       # number of replicates
    )

  lod_simulation <- tibble::tibble(
    x = exp_model$time,
    y = exp_model$random_y * sample(1:1000000, 1),
    `Sample Name` = as.character(exp_model$time))

  return(lod_simulation)

}

lod_simulated_database <-
  tibble::tibble(
    ID = paste("Limit Of Detection", 
               stringr::str_pad(1:200, width = 3, pad = "0")),
    `Dilution Batch Name` = "Limit Of Detection"

  ) |> 
  dplyr::mutate(
    dilution_data = purrr::map(
      .x = .data$ID,
      .f = simulate_lod
      )
    ) |> 
  tidyr::unnest("dilution_data")

```

```{r}

# Create dilution statistical summary
saturation_dilution_summary <- summarise_dilution_table(
  saturation_simulated_database,
  grouping_variable = c("ID"),
  conc_var = "x",
  signal_var = "y")
```

```{r}

lod_dilution_summary <- summarise_dilution_table(
  lod_simulated_database,
  grouping_variable = c("ID"),
  conc_var = "x",
  signal_var = "y")

```

```{r}
saturation_ggplot_table <- add_ggplot_panel(
  dilution_table = saturation_simulated_database,
  dilution_summary = saturation_dilution_summary,
  grouping_variable = c("ID"),
  dil_batch_var = "Dilution Batch Name",
  conc_var = "x",
  signal_var = "y",
  conc_var_units = ""
  )

saturation_ggplot_list <- saturation_ggplot_table$panel

view_ggplot_pdf(saturation_ggplot_list,
                filename = "dilution_plot_saturated.pdf",
                ncol_per_page = 2,
                nrow_per_page = 2)

```

```{r}
lod_ggplot_table <- add_ggplot_panel(
  dilution_table = lod_simulated_database,
  dilution_summary = lod_dilution_summary,
  grouping_variable = c("ID"),
  dil_batch_var = "Dilution Batch Name",
  conc_var = "x",
  signal_var = "y",
  conc_var_units = ""
  )

lod_ggplot_list <- lod_ggplot_table$panel

view_ggplot_pdf(lod_ggplot_list,
                filename = "dilution_plot_lod.pdf",
                ncol_per_page = 2,
                nrow_per_page = 2)

```

```{r}
# Create a trellis table
saturation_trellis_table <- add_plotly_panel(
  dilution_table = saturation_simulated_database,
  dilution_summary = saturation_dilution_summary,
  grouping_variable = c("ID"),
  sample_name_var = "Sample Name",
  dil_batch_var = "Dilution Batch Name",
  dil_batch_col = c(
    "#377eb8"),
  conc_var = "x",
  conc_var_units = "",
  conc_var_interval = 50,
  signal_var = "y",
  have_plot_title = FALSE
) |> 
  convert_to_cog(
    grouping_variable = "ID",
    panel_variable = "panel",
    col_name_vec = "col_name_vec",
    desc_vec = "desc_vec",
    type_vec = "type_vec"
  )

view_trellis_html(
  trellis_table = saturation_trellis_table,
  grouping_variable = c("ID"),
  trellis_report_name = "Dilution_Plot_Saturated",
  trellis_report_folder = "Dilution_Saturated_Folder",
  panel_variable = "panel"
)
```

```{r}
# Create a trellis table
lod_trellis_table <- add_plotly_panel(
  dilution_table = lod_simulated_database,
  dilution_summary = lod_dilution_summary,
  grouping_variable = c("ID"),
  sample_name_var = "Sample Name",
  dil_batch_var = "Dilution Batch Name",
  dil_batch_col = c(
    "#377eb8"),
  conc_var = "x",
  conc_var_units = "",
  conc_var_interval = 50,
  signal_var = "y",
  have_plot_title = FALSE
) |> 
  convert_to_cog(
    grouping_variable = "ID",
    panel_variable = "panel",
    col_name_vec = "col_name_vec",
    desc_vec = "desc_vec",
    type_vec = "type_vec"
  )

view_trellis_html(
  trellis_table = lod_trellis_table,
  grouping_variable = c("ID"),
  trellis_report_name = "Dilution_Plot_LOD",
  trellis_report_folder = "Dilution_LOD_Folder",
  panel_variable = "panel"
)
```