### 2023-01-27 README.qmd
Run code with new version of `lancer`.

---
title: "`lancer` Output Examples"
format: gfm
---

[![](https://img.shields.io/badge/doi-10.6084/m9.figshare.21965324-blue.svg?style=flat&labelColor=whitesmoke&logo=data%253Aimage%252Fpng%253Bbase64%252CiVBORw0KGgoAAAANSUhEUgAAAB8AAAAfCAYAAAAfrhY5AAAJsklEQVR42qWXd1DTaRrHf%252BiB2Hdt5zhrAUKz4IKEYu9IGiGFFJJQ0gkJCAKiWFDWBRdFhCQUF3UVdeVcRQEBxUI3yY9iEnQHb3bdW1fPubnyz%252F11M7lvEHfOQee2ZOYzPyDv%252B3yf9%252Fk95YX4fx%252BltfUt08GcFEuPR4U9hDDZ%252FVngIlhb%252FSiI6InkTgLzgDcgfvtnovhH4BzoVlrbwr55QnhCtBW4QHXnFrZbPBaQoBh4%252FSYH2EnpBEtqcDMVzB93wA%252F8AFwa23XFGcc8CkT3mxz%252BfXWtq9T9IQlLIXYEuHojudb%252BCM7Hgdq8ydi%252FAHiBXyY%252BLjwFlAEnS6Jnar%252FvnQVhvdzasad0eKvWZKe8hvDB2ofLZ%252FZEcWsh%252BhyIuyO5Bxs2iZIE4nRv7NWAb0EO8AC%252FWPxjYAWuOEX2MSXZVgPxzmRL3xKz3ScGpx6p6QnOx4mDIFqO0w6Q4fEhO5IzwxlSwyD2FYHzwAW%252BAZ4fEsf74gCumykwNHskLM7taQxLYjjIyy8MUtraGhTWdkfhkFJqtvuVl%252F9l2ZquDfEyrH8B0W06nnpH3JtIyRGpH1iJ6SfxDIHjRXHJmdQjLpfHeN54gnfFx4W9QRnovx%252FN20aXZeTD2J84hn3%252BqoF2Tqr14VqTPUCIcP%252B5%252Fly4qC%252BUL3sYxSvNj1NwsVYPsWdMUfomsdkYm3Tj0nbV0N1wRKwFe1MgKACDIBdMAhPE%252FwicwNWxll8Ag40w%252BFfhibJkGHmutjYeQ8gVlaN%252BjO51nDysa9TwNUFMqaGbKdRJZFfOJSp6mkRKsv0rRIpEVWjAvyFkxNOEpwvcAVPfEe%252Bl8ojeNTx3nXLBcWRrYGxSRjDEk0VlpxYrbe1ZmaQ5xuT0u3r%252B2qe5j0J5uytiZPGsRL2Jm32AldpxPUNJ3jmmsN4x62z1cXrbedXBQf2yvIFCeZrtyicZZG2U2nrrBJzYorI2EXLrvTfCSB43s41PKEvbZDEfQby6L4JTj%252FfIwam%252B4%252BwucBu%252BDgNK05Nle1rSt9HvR%252FKPC4U6LTfvUIaip1mjIa8fPzykii23h2eanT57zQ7fsyYH5QjywwlooAUcAdOh5QumgTHx6aAO7%252FL52eaQNEShrxfhL6albEDmfhGflrsT4tps8gTHNOJbeDeBlt0WJWDHSgxs6cW6lQqyg1FpD5ZVDfhn1HYFF1y4Eiaqa18pQf3zzYMBhcanlBjYfgWNayAf%252FASOgklu8bmgD7hADrk4cRlOL7NSOewEcbqSmaivT33QuFdHXj5sdvjlN5yMDrAECmdgDWG2L8P%252BAKLs9ZLZ7dJda%252BB4Xl84t7QvnKfvpXJv9obz2KgK8dXyqISyV0sXGZ0U47hOA%252FAiigbEMECJxC9aoKp86re5O5prxOlHkcksutSQJzxZRlPZmrOKhsQBF5zEZKybUC0vVjG8PqOnhOq46qyDTDnj5gZBriWCk4DvXrudQnXQmnXblebhAC2cCB6zIbM4PYgGl0elPSgIf3iFEA21aLdHYLHUQuVkpgi02SxFdrG862Y8ymYGMvXDzUmiX8DS5vKZyZlGmsSgQqfLub5RyLNS4zfDiZc9Edzh%252FtCE%252BX8j9k%252FqWB071rcZyMImne1SLkL4GRw4UPHMV3jjwEYpPG5uW5fAEot0aTSJnsGAwHJi2nvF1Y5OIqWziVCQd5NT7t6Q8guOSpgS%252Fa1dSRn8JGGaCD3BPXDyQRG4Bqhu8XrgAp0yy8DMSvvyVXDgJcJTcr1wQ2BvFKf65jqhvmxXUuDpGBlRvV36XvGjQzLi8KAKT2lYOnmxQPGorURSV0NhyTIuIyqOmKTMhQ%252BieEsgOgpc4KBbfDM4B3SIgFljvfHF6cef7qpyLBXAiQcXvg5l3Iunp%252FWv4dH6qFziO%252BL9PbrimQ9RY6MQphEfGUpOmma7KkGzuS8sPUFnCtIYcKCaI9EXo4HlQLgGrBjbiK5EqMj2AKWt9QWcIFMtnVvQVDQV9lXJJqdPVtUQpbh6gCI2Ov1nvZts7yYdsnvRgxiWFOtNJcOMVLn1vgptVi6qrNiFOfEjHCDB3J%252BHDLqUB77YgQGwX%252Fb1eYna3hGKdlqJKIyiE4nSbV8VFgxmxR4b5mVkkeUhMgs5YTi4ja2XZ009xJRHdkfwMi%252BfocaancuO7h%252FMlcLOa0V%252FSw6Dq47CumRQAKhgbOP8t%252BMTjuxjJGhXCY6XpmDDFqWlVYbQ1aDJ5Cptdw4oLbf3Ck%252BdWkVP0LpH7s9XLPXI%252FQX8ws%252Bj2In63IcRvOOo%252BTTjiN%252BlssfRsanW%252B3REVKoavBOAPTXABW4AL7e4NygHdpAKBscmlDh9Jysp4wxbnUNna3L3xBvyE1jyrGIkUHaqQMuxhHElV6oj1picvgL1QEuS5PyZTEaivqh5vUCKJqOuIgPFGESns8kyFk7%252FDxyima3cYxi%252FYOQCj%252F%252B9Ms2Ll%252Bhn4FmKnl7JkGXQGDKDAz9rUGL1TIlBpuJr9Be2JjK6qPzyDg495UxXYF7JY1qKimw9jWjF0iV6DRIqE%252B%252FeWG0J2ofmZTk0mLYVd4GLiFCOoKR0Cg727tWq981InYynvCuKW43aXgEjofVbxIqrm0VL76zlH3gQzWP3R3Bv9oXxclrlO7VVtgBRpSP4hMFWJ8BrUSBCJXC07l40X4jWuvtc42ofNCxtlX2JH6bdeojXgTh5TxOBKEyY5wvBE%252BACh8BtOPNPkApjoxi5h%252B%252FFMQQNpWvZaMH7MKFu5Ax8HoCQdmGkJrtnOiLHwD3uS5y8%252F2xTSDrE%252F4PT1yqtt6vGe8ldMBVMEPd6KwqiYECHDlfbvzphcWP%252BJiZuL5swoWQYlS%252Br7Yu5mNUiGD2retxBi9fl6RDGn4Ti9B1oyYy%252BMP5G87D%252FCpRlvdnuy0PY6RC8BzTA40NXqckQ9TaOUDywkYsudxJzPgyDoAWn%252BB6nEFbaVxxC6UXjJiuDkW9TWq7uRBOJocky9iMfUhGpv%252FdQuVVIuGjYqACbXf8aa%252BPeYNIHZsM7l4s5gAQuUAzRUoT51hnH3EWofXf2vkD5HJJ33vwE%252FaEWp36GHr6GpMaH4AAPuqM5eabH%252FhfG9zcCz4nN6cPinuAw6IHwtvyB%252FdO1toZciBaPh25U0ducR2PI3Zl7mokyLWKkSnEDOg1x5fCsJE9EKhH7HwFNhWMGMS7%252BqxyYsbHHRUDUH4I%252FAheQY7wujJNnFUH4KdCju83riuQeHU9WEqNzjsJFuF%252FdTDAZ%252FK7%252F1WaAU%252BAWymT59pVMT4g2AxcwNa0XEBDdBDpAPvgDIH73R25teeuAF5ime2Ul0OUIiG4GpSAEJeYW9wDTf43wfwHgHLKJoPznkwAAAABJRU5ErkJggg%253D%253D)](https://doi.org/10.6084/m9.figshare.21965324) [![](https://img.shields.io/badge/License-CC%20BY%204.0-yellow.svg)](https://creativecommons.org/licenses/by/4.0/)

# Introduction

Here is where I store some output example of the R package [`lancer`](https://github.com/SLINGhub/lancer).

# Version control

Refer to [NEWS.md](https://github.com/JauntyJJS/lancer_output/blob/main/NEWS.md) for more information.

# Output Examples

## Excel Output Example

[`curve_summary.xlsx`](https://raw.githubusercontent.com/JauntyJJS/lancer_output/main/curve_summary.xlsx) is an example of `lancer`'s output in Excel.

## Pdf Output Example

[`curve_plot.pdf`](https://raw.githubusercontent.com/JauntyJJS/lancer_output/main/curve_plot.pdf) is an example of `lancer`'s output in pdf.

## Trellis Output Example

https://lancer-interactive-example.netlify.app is an example of `lancer`'s output as an interactive trellis plot from the folder `Quarto_Example/Curve_Plot_Folder` when the `index.html` file inside the folder is clicked.

## Quarto Output Example

https://lancer-quarto-example.netlify.app comes from left clicking the `index.html` file from the `Quarto_Example` folder. The `index.html` file is generated by running [Quarto](https://quarto.org/) on the `index.qmd` script file.

## Simulation Report

https://lancer-simulation.netlify.app is the report of a simulation experiment done in the paper. The link comes from left clicking the `index.html` file from the `Quarto_Simulation` folder. The `index.html` file is generated by running [Quarto](https://quarto.org/) on the `index.qmd` script file.

# Setting up

In order to use this project locally, install

-   [R](https://cran.rstudio.com/)
-   [RStudio](https://posit.co/download/rstudio-desktop/)
-   [git](https://git-scm.com/)
-   [GNU software development tools](https://support.posit.co/hc/en-us/articles/200486498-Package-Development-Prerequisites) like
    -   [RTools](https://cran.rstudio.com/bin/windows/Rtools/) for Windows

    -   [Xcode](http://itunes.apple.com/us/app/xcode/id497799835?mt=12) for Mac

    -   `r-base-dev` for Linux

[Clone the repository](https://happygitwithr.com/existing-github-first.html#new-rstudio-project-via-git-clone), then run the following code chunks to initialise the project.

First, install the R package [`renv`](https://rstudio.github.io/renv/index.html)

```{r}
#| eval: false
#| warning: false
#| message: false

install.packages("renv", dependencies = TRUE)
```

Once installation is done, start a local R library.

```{r}
#| eval: false
#| warning: false
#| message: false

renv::init(bare = TRUE)
```

Some of the R packages may only have the source code available, without any binary files. Install these R packages to ensure installation of such R packages is possible.

```{r}
#| eval: false
#| warning: false
#| message: false

install.packages("rlang", dependencies = TRUE)
install.packages("devtools", dependencies = TRUE)
```

Type the following command to install R packages `lancer` version 0.1.0.

```{r}
#| eval: false
#| warning: false
#| message: false
devtools::install_github("SLINGhub/lancer", ref = "0.1.0")
```

Next, type the following command to install the other R packages dependencies.

```{r}
#| eval: false
#| warning: false
#| message: false
renv::restore()
```

Once you have run these steps, in RStudio you can open the individual Quarto documents and render each file.

# Quarto documents

## Quarto_Example folder

Rendering `index.qmd` will produce in the folder

-    `curve_summary.xlsx`
-    `curve_plot.pdf`
-    `index.html` with dependencies `index_files` and `Curve_Plot_Folder`

## Quarto_Simulation folder

Rendering `index.qmd` will produce in the folder

-    `linear_simulated_database.csv`, `noise_regime_simulated_database.csv` and `saturation_regime_simulated_database.csv`
-    `linear_simulated_database.rds`, `noise_regime_simulated_database.rds` and `saturation_regime_simulated_database.rds`
-    `linear_curve_summary.csv`, `noise_regime_curve_summary.csv` and `saturation_regime_curve_summary.csv`
-    `linear_curve_summary.rds`, `noise_regime_curve_summary.rds` and `saturation_regime_curve_summary.rds`
-    `curve_plot_linear.pdf`, `curve_plot_noise_regime.pdf` and `curve_plot_saturation_regime.pdf`
-    `index.html` with dependencies `index_files`, `Curve_Linear_Folder`, `Curve_Noise_Regime_Folder` and `Curve_Saturation_Regime_Folder`

## Manuscript folder

Rendering `manuscript.qmd` will produce `manuscript.docx`, the manuscript document.

Rendering `supplementary.qmd` will produce `supplementary.docx`, the supplementary document.

Do note that both Quarto documents read in the results of the simulated datasets from the files `linear_curve_summary.rds`, `noise_regime_curve_summary.rds` and `saturation_regime_curve_summary.rds` in the `Quarto_Simulation folder`